---
title: "Population Dinamics"
author: "Elena Pazhenkova"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr) #try it later
library(ggplot2)
```

## Initialization 
Upload data with initial population size and propability ranges of fecundity (F) and survival rate (S).
Then set up number of cycles and runs.
Create empty dataframe to save population size of each age class per cycle.

```{r import_and_set_up}
data <- read.csv('/Users/irina/Desktop/input.csv', header = F, sep=" ", stringsAsFactors = FALSE)
cycles = 30
runs = 500
output_1 <- cbind(matrix(0, nrow = 1, ncol =1),t(data[,2])) #dataframe for population size in each age class
```

## Functions

We observe how our population size changes during 30 cycles.
This is the  **run** function. In each run the new Leslie matrix with main parameters is created.
```{r}
run <- function(){
Leslie_matrix <-  matrix(0, nrow(data), nrow(data)) #new Leslie matrix
Leslie_matrix[1,(3:6)] <- sapply(data[3:6, 3], function(x) runif(1, min = as.numeric(strsplit(x, "-")[[1]][1]), max = as.numeric(strsplit(x, "-")[[1]][2]))) #Initialization of fecundity for each age in this run
diag(Leslie_matrix[2:nrow(Leslie_matrix), 1:ncol(Leslie_matrix)-1]) <- sapply(data[1:5, 4], function(x) runif(1, min = as.numeric(strsplit(x, "-")[[1]][1]), max = as.numeric(strsplit(x, "-")[[1]][2]))) #Initialization of survival rate for each age in this run
initial_n <- data[,2]
for (n in 1:cycles) {
  initial_n <- Leslie_matrix%*%initial_n #compute population size for each age class
  output_1 <<- rbind(output_1, c(n, initial_n))  #Write population size for each age class in this run
  }
return(c(Leslie_matrix[1,(3:6)], diag(Leslie_matrix[2:nrow(Leslie_matrix), 1:ncol(Leslie_matrix)-1]))) #Save parameters of the run
}
```


## Output

```{r}
output_2 <-  t(data.frame(replicate(runs, run()))) #Create dataframe with parameters of each run
colnames(output_2) <- c(paste0("F", 3:6), paste0("S", 1:5))
write.csv(output_2)
output_1 <- tibble::rownames_to_column(data.frame(output_1))
colnames(output_1) <-  c("run", "cycle", 0:5)
write.csv(output_1)
```

##Visualization

You can also embed plots, for example:

```{r plot, echo=FALSE}
head(output_1)
nrow(output_1)
ggplot(output1, aes())


output_1[2:8] %>% tidyr::gather("generation", "value", 2:7) %>% 
  ggplot(., aes(cycle, value, color = generation ))+
  geom_point(size = 0.4, alpha = 0.7)+
  geom_smooth(method = "lm", se=T, color = "darkgray", size = 0.5)+
  facet_wrap(~generation)


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
